Faithful Variable Screening for High-Dimensional Convex Regression

### 2016-03-08 tex/scam/faithful.tex
edits for final version

\documentclass[12pt,pdftex,aos]{imsart}
%\documentclass[12pt,pdftex,generic,noinfoline]{imsart}

\RequirePackage[OT1]{fontenc}
\usepackage{amsthm,amsmath,amsfonts,natbib,mathtools,custom_math}
\RequirePackage{hypernat}
\usepackage[ruled,section]{algorithm}
\usepackage[noend]{algorithmic}
\usepackage{graphicx}
\usepackage[hscale=0.7,vscale=0.8]{geometry}
\usepackage{subfigure}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{yhmath}

\renewcommand{\baselinestretch}{1.1}
%\parskip12pt
%\parindent0pt
\setcounter{tocdepth}{2}

@@

\begin{aug}
\vskip10pt
\author{\fnms{Min} \snm{Xu}\ead[label=e1]{minx@cs.cmu.edu}}
\comma 
\author{\fnms{Minhua} \snm{Chen}\ead[label=e2]{minhua@galton.uchicago.edu}}
\and
\author{\fnms{John} \snm{Lafferty}\ead[label=e3]{lafferty@galton.uchicago.edu}}
\address{
\begin{tabular}{ccc}
\\[10pt]
Machine Learning Department & Department of Statistics & Department of Statistics\\
Carnegie Mellon University  & University of Chicago & Department of Computer Science\\
& & University of Chicago
\end{tabular}
%\\[20pt]
%\printead{e1,e2,e3}
}
\end{aug}

\input{abstract}
%\tableofcontents
\vskip20pt
\end{frontmatter}

\maketitle
\vskip10pt

%\tableofcontents

@@

\clearpage

\newpage
\input{appendix}
%\newpage
%\input{notes}

\section*{Acknowledgements}
Research supported in part by NSF grants IIS-1116730, 
AFOSR grant FA9550-09-1-0373, ONR grant
N000141210762, and an Amazon AWS in Education Machine Learning
Research grant.

\bibliographystyle{imsart-nameyear}
\bibliography{local}

>>>>

\documentclass[pdftex,aos]{imsart}

\RequirePackage[OT1]{fontenc}
\RequirePackage{amsthm,amsmath}
\usepackage{amsfonts,mathtools,custom_math,amssymb}
\RequirePackage[numbers]{natbib}
\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
\usepackage{subfigure}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage{multirow}
\usepackage{xr}
\externaldocument{supplement}

% settings
%\pubyear{2005}
%\volume{0}
%\issue{0}
%\firstpage{1}
%\lastpage{8}
\arxiv{arXiv:0000.0000}

\startlocaldefs
\numberwithin{equation}{section}
\theoremstyle{plain}
\newtheorem{thm}{Theorem}[section]
\endlocaldefs

@@

\begin{aug}
\author{\fnms{Min} \snm{Xu}\thanksref{t1,m1}\ead[label=e1]{min.cut@gmail.com}},
\author{\fnms{Minhua} \snm{Chen}\thanksref{m2}\ead[label=e2]{minhua@somewhere.com}}
\and
\author{\fnms{John} \snm{Lafferty}\thanksref{t1,m3}\ead[label=e3]{lafferty@galton.uchicago.edu}}

\thankstext{t1}{Research supported in part by NSF grants IIS-1116730 and DMS-1513594,
AFOSR grant FA9550-09-1-0373, ONR grant N00014-12-10762, and an Amazon
AWS in Education Machine Learning Research grant.}
\runauthor{M.~Xu, M.~Chen and J.~Lafferty}

\affiliation{University of Pennsylvania\thanksmark{m1}, Amazon.com\thanksmark{m2}, and University of Chicago\thanksmark{m3}}

\address{Address of M.~Xu\\
Department of Statistics\\
The Wharton School\\
University of Pennsylvania\\
Philadelphia, Pennsylvania 19104\\
USA\\
\printead{e1}\\
}

\address{Address of M.~Chen\\
Amazon.com\\
410 Terry Avenue North\\
Seattle, Washington 98109\\
USA\\
\printead{e2}\\}

\address{Address of J.~Lafferty\\
Department of Statistics\\
University of Chicago\\
5734 South University Avenue\\
Chicago, Illinois 60637\\
USA\\
\printead{e3}\\}

\end{aug}

\begin{abstract}
  We study the problem of variable selection in convex nonparametric
  regression.  Under the assumption that the true regression function
  is convex and sparse, we develop a screening procedure to select a
  subset of variables that contains the relevant variables. Our
  approach is a two-stage quadratic programming method that estimates
  a sum of one-dimensional convex functions, followed by
  one-dimensional concave regression fits on the residuals.  In
  contrast to previous methods for sparse additive models, the
  optimization is finite dimensional and requires no tuning parameters
  for smoothness. Under appropriate assumptions, we prove that the
  procedure is faithful in the population setting, yielding no false
  negatives.  We give a finite sample statistical analysis, and
  introduce algorithms for efficiently carrying out the required
  quadratic programs. The approach leads to computational and
  statistical advantages over fitting a full model, and provides an
  effective, practical approach to variable screening in convex
  regression.
\end{abstract}

\begin{keyword}[class=MSC]
\kwd[Primary ]{62G08}
\kwd[; secondary ]{52A41}
\end{keyword}

\begin{keyword}
\kwd{nonparametric regression}
\kwd{convex regression}
\kwd{variable selection}
\kwd{quadratic programming}
\kwd{additive model}
\end{keyword}

\end{frontmatter}

@@

\section*{Acknowledgements}
The authors thank  Ming Yuan for valuable discussions and suggestions
on this research. The authors are also grateful to the anonymous
reviewers whose careful proofreading and helpful comments significantly
improved the quality of the paper.

\begin{supplement}[id=suppA]
%\sname{Supplement A}
\stitle{Supplement to Faithful Variable Screening for High-Dimensional Convex Regression}
\slink[doi]{doi}
\sdatatype{.pdf}
\sdescription{The supplement provides detailed proofs of certain technical results,
together with further explanation of the Gaussian example
and simplifications when the density is a product.}
\end{supplement}

%\clearpage
\bibliographystyle{imsart-nameyear}
\bibliography{local}

### 2016-03-08 tex/scam/introduction.tex
mods for supplement
edits for final version

Recently, there has been increased research activity on shape-constrained estimation. \cite{guntusen:13} analyze univariate convex regression and show surprisingly that the risk of the MLE is adaptive to the complexity of the true function. \cite{seijo2011nonparametric} and \cite{Lim:12} study maximum likelihood estimation of multivariate convex regression and independently establish its consistency. \cite{Cule:10} and \cite{kim2014global} analyze log-concave density estimation and prove consistency of the MLE; the latter further show that log-concave density estimation has minimax risk lower bounded by $n^{-2/(d+1)}$ for $d \geq 2$, refuting a common notion that the condition of convexity is equivalent, in estimation difficulty, to the condition of having two bounded derivatives. Additive shape-constrained estimation has also been studied; \cite{pya2014shape} propose a penalized B-spline estimator while \cite{chen2014generalised} show the consistency of the MLE. To the best of our knowledge however, there has been no work on variable selection and estimation of high-dimensional convex functions.

@@

Perhaps more closely related to the present work is the framework
studied by \cite{Raskutti:12} for sparse additive models, where sparse
regression is considered under an additive assumption, with each
component function belonging to an RKHS.  An advantage of working over
an RKHS is that nonparametric regression with a sparsity-inducing
regularization penalty can be formulated as a finite dimensional
convex cone optimization.  On the other hand, smoothing parameters for
the component Hilbert spaces must be chosen, leading to extra tuning
parameters that are difficult to select in practice. There has also been
work on estimating sparse additive models over a spline basis, for 
instance the work of \cite{huang2010variable}, but these approaches also
require the tuning of smoothing parameters.

@@

In the following section we give a high-level summary of our technical
results, including additive faithfulness, variable selection 
consistency, and high dimensional scaling.  In
Section~\ref{sec:additivefaithful} we give a detailed account
of our method and the conditions under which we can guarantee
consistent variable selection.  In Section~\ref{sec:optimization}
we show how the required quadratic programs can be reformulated
to be more efficient and scalable.  In Section~\ref{sec:finitesample}
we give the details of our finite sample analysis, showing
that a sample size growing as $n = O\big(\textrm{poly}(s) \log p\big)$
is sufficient for variable selection.  In Section~\ref{sec:thesims}
we report the results of simulations that illustrate our methods
and theory.  The full proofs are
given in a technical appendix.

>>>>

Recently, there has been increased research activity on
shape-constrained estimation. \citet{guntusen:13} analyze univariate
convex regression and show surprisingly that the risk of the MLE is
adaptive to the complexity of the true
function. \citet{seijo2011nonparametric} and \citet{Lim:12} study
maximum likelihood estimation of multivariate convex regression and
independently establish its consistency. \citet{Cule:10} and
\citet{kim2014global} analyze log-concave density estimation and prove
consistency of the MLE; the latter further show that log-concave
density estimation has minimax risk lower bounded by $n^{-2/(d+1)}$
for $d \geq 2$, refuting a common notion that the condition of
convexity is equivalent, in estimation difficulty, to the condition of
having two bounded derivatives. Additive shape-constrained estimation
has also been studied; \citet{pya2014shape} propose a penalized
B-spline estimator while \citet{chen2014generalised} show the
consistency of the MLE. To the best of our knowledge however, there
has been no work on variable selection and estimation of
high-dimensional convex functions.

@@

Perhaps more closely related to the present work is the framework
studied by \citet{Raskutti:12} for sparse additive models, where sparse
regression is considered under an additive assumption, with each
component function belonging to an RKHS.  An advantage of working over
an RKHS is that nonparametric regression with a sparsity-inducing
regularization penalty can be formulated as a finite dimensional
convex cone optimization.  On the other hand, smoothing parameters for
the component Hilbert spaces must be chosen, leading to extra tuning
parameters that are difficult to select in practice. There has also been
work on estimating sparse additive models over a spline basis, for 
instance the work of \cite{huang2010variable}, but these approaches also
require the tuning of smoothing parameters.

@@

In the following section we give a high-level summary of our technical
results, including additive faithfulness, variable selection 
consistency, and high dimensional scaling.  In
Section~\ref{sec:additivefaithful} we give a detailed account
of our method and the conditions under which we can guarantee
consistent variable selection.  In Section~\ref{sec:optimization}
we show how the required quadratic programs can be reformulated
to be more efficient and scalable.  In Section~\ref{sec:finitesample}
we give the details of our finite sample analysis, showing
that a sample size growing as $n = O\big(\textrm{poly}(s) \log p\big)$
is sufficient for variable selection.  In Section~\ref{sec:thesims}
we report the results of simulations that illustrate our methods
and theory.  The full proofs are
given in the supplementary material \citep{supplement}.
